Dataset Summary

### 2019-11-06 Prep & old/prep and old- scrrpts/2018.07.03-meta_analysis_prototype.Rmd
rearrange

---
title: "Protoype for metaanalysis"
author: "Daniel Falster & Alistair Senior + Shinichi messing things up"
date: "03/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: show
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: "Protoype for metaanalysis"
author: "Daniel Falster & Alistair Senior"
date: "03/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: show
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

### 2019-11-06 Prep & old/prep and old- scrrpts/2018.08.28-Overview.Rmd
rearrange

---
title: "Dataset Summary"
author: "Daniel Falster & Susie Zajitschek"
date: "06/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE)
root.dir = rprojroot::find_root("README.md")
knitr::opts_knit$set(root.dir = root.dir)

library(readr)
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
library(viridis)
library(knitr)
library(kableExtra)
```

Here is an initial exploration of the mouse data we are working on.

The data is big, but not so large that we can't use our standard tools on a desktop. I'd suggest using packages from the [tidyverse](https://www.tidyverse.org/) family, in particular `readr`, `dplyr`, `ggplot2`, `skimr`, `scales`, `viridis`.

#  Dataset overview

Read in data, specifying variable types:

```{r cars}
data_raw <- read_csv("data/dr7.0_all_control_data.csv", 
                  col_types = cols(
                      .default = col_character(),
                      project_id = col_character(),
                      id = col_character(),
                      parameter_id = col_character(),
                      age_in_days = col_integer(),
                      date_of_experiment = col_datetime(format = ""),
                      weight = col_double(),
                      phenotyping_center_id = col_character(),
                      production_center_id = col_character(),
                      weight_date = col_datetime(format = ""),
                      date_of_birth = col_datetime(format = ""),
                      procedure_id = col_character(),
                      pipeline_id = col_character(),
                      biological_sample_id = col_character(),
                      biological_model_id = col_character(),
                      weight_days_old = col_integer(),
                      datasource_id = col_character(),
                      experiment_id = col_character(),
                      data_point = col_double(),
                      age_in_weeks = col_integer(),
                      `_version_` = col_character()
                      )
                )
```

@@

```{r}
data_raw %>% dim()
```

Of the `r data %>% names() %>% length()` variables in the dataset, lots of variables are full of NAs:

```{r}
n_records <- data_raw %>% nrow()
NA_count <- data_raw %>%
  summarise_all(funs(sum(is.na(.))/n_records*100)) %>% 
  unlist() %>% sort(decreasing=TRUE) %>%
  tibble(variable=names(.), percent_NA = .)

# make a plot
ggplot(NA_count, aes(percent_NA)) + geom_histogram(bins=50)
```

Let's remove those variables that have all NAs:

```{r}
(all_NAs <- NA_count$variable[NA_count$percent_NA == 100])
data <- data_raw %>% select(-one_of(all_NAs))
```

@@

@@

```{r}
select(filter(data_raw, biological_sample_id == "107609"), sex, production_center, external_sample_id, biological_sample_id, age_in_days, weight, parameter_name) %>% arrange(age_in_days) %>% data.frame()
```

```{r}

data_i <- data %>% 
  filter(parameter_name == "Body weight") %>%
  group_by(biological_sample_id) %>% 
  summarise(
    production_center = production_center[1],
    project_name = project_name[1],
    sex = sex[1],
    n = n(),
    med_age= median(age_in_days)
    )

data_i
```

# Datasource

more recently, Jeremy states: "I provided you too much data (!) which should not be included.  A quarter to a third of the data is from legacy projects and should be excluded.  As I mentioned, exclude anything without datasource_name = `IMPC`"

By source: 
```
data_raw$datasource_name %>% table()
```

So reducing to IMPC reduces the data by one third.

It also greatly reduces the number of parameters:
```{r}
filter(data_raw, datasource_name == 'IMPC') %>% pull(parameter_name) %>% unique() %>% length()
```

```
data_raw %>% pull(parameter_name) %>% unique() %>% length()
```

# Issues:

1.  `procedure_name` are values eding in `(GMC)` different?
2.  Spelling / case differences: e.g. `Body weight` vs `Body Weight`

>>>>

---
title: "Dataset Summary"
author: "Daniel Falster & Susie Zajitschek"
date: "06/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: show
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE)
root.dir = rprojroot::find_root("README.md")
knitr::opts_knit$set(root.dir = root.dir)
```

```{r}
library(readr)
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
library(viridis)
library(knitr)
library(pander)
library(kableExtra)
source("R/data_load_clean.R")
```

# Loading data

Read in cleaned data, specifying variable types:

```{r}
data <- readRDS("export/data_clean.rds")
```

Note: cleaned data were generated by running

```{r, eval=FALSE}
data <- load_raw("data/dr7.0_all_control_data.csv") %>% clean_raw_data()
```

where using the follwing function
```{r}
clean_raw_data
```

# Data overview

@@

```{r}
data %>% dim()
```

@@

# Strains

There are several strains under the variable `strain_name`:

```{r}
data$strain_name %>% unique() %>% length()
```

```{r}
data$strain_name %>% table() %>% sort(decreasing = TRUE) %>%
  tibble(variable=names(.), count = .) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

There is also a variable  several strains under the variable `strain_accession_id`:

```{r}
data$strain_accession_id %>% unique() %>% length()
```

```{r}
data$strain_accession_id %>% table() %>% sort(decreasing = TRUE) %>%
  tibble(variable=names(.), count = .) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

@@

```{r}
select(filter(data, biological_sample_id == "107609"), sex, production_center, biological_sample_id, age_in_days, weight, parameter_name) %>% arrange(age_in_days) %>% data.frame()
```
